Supplementary Information

### 2020-11-06 doc/manuscript/MEE_manuscript.Rmd
Update MEE_manuscript.Rmd
#84

**We used the *climwin* package in R [@van_de_pol_identifying_2016] to identify the most important climate driver and the time window over which its effect was strongest for each of two categories of variables: a temperature group (mean, min, and max temperature; PET) and a precipitation group (precipitation, number of days with precipitation).**
To remove low-frequency variation that most likely represents responses to non-climatic drivers (*e.g.*, growth and aging of the tree, change in competitive dynamics, atmospheric pollution), we detrended the response variables by using a generalized additive model (GAM) to fit a spline to individual growth records ($RW$, $BAI$, or $\Delta AGB$) from each core, thereby producing residuals.
We then used *climwin* to identify the climate variables most strongly correlated to the residuals of $RW$, $BAI$, or $\Delta AGB$, specifying quadratic fits to allow for potential nonlinearities in the climate response.
Within *climwin*, we specified a mixed effects model using species and core identity as random effects: `residual growth index ~ [climate] + (1 | sp) + (1 | treeID)`. 
Here, for each permutation, `climate` specifies one of the climate drivers in the climate variable group, analyzed over one of all possible combinations of consecutive months over a 15 month period ending near the time of cessation of formation of each annual ring (*Table S1*). (**[ISSUE #51 in ForestGEO-climate-sensitivity ](https://github.com/EcoClimLab/ForestGEO-climate-sensitivity/issues/51)**)
*Climwin* runs all potential models to select the best fit (lowest AIC), and does k-fold cross-validation in its computation of AIC to guard against over-fitting [@van_de_pol_identifying_2016]. 
For each group of candidate climate variables, we move forward with the best variable over the time window identified by *climwin* as a candidate climate variable for the multivariate models.

>>>>

**We used the *climwin* package in R [@van_de_pol_identifying_2016] to identify the most important climate driver and the time window over which its effect was strongest for each of two categories of variables: a temperature group (mean, min, and max temperature; PET) and a precipitation group (precipitation, number of days with precipitation).**
To remove low-frequency variation that most likely represents responses to non-climatic drivers (*e.g.*, growth and aging of the tree, change in competitive dynamics, atmospheric pollution), we detrended the response variables by using a generalized additive model (GAM) to fit a spline to individual growth records ($RW$, $BAI$, or $\Delta AGB$) from each core, thereby producing residuals.
We then used *climwin* to identify the climate variables most strongly correlated to the residuals of $RW$, $BAI$, or $\Delta AGB$, specifying quadratic fits to allow for potential nonlinearities in the climate response.
Within *climwin*, we specified a mixed effects model using species and core identity as random effects: `residual growth index ~ [climate] + (1 | sp) + (1 | treeID)`.
We allowed first-order linear or concave-down second-order polynomial fits, but ignored any concave-up fits on the basis that these are not expected biologically and, when identified in initial analyses, often appeared to be cases of over-fitting (**[ISSUE #84 in ForestGEO-climate-sensitivity ](https://github.com/EcoClimLab/ForestGEO-climate-sensitivity/issues/84)**).

Here, for each permutation, `climate` specifies one of the climate drivers in the climate variable group, analyzed over one of all possible combinations of consecutive months over a 15 month period ending near the time of cessation of formation of each annual ring (*Table S1*). (**[ISSUE #51 in ForestGEO-climate-sensitivity ](https://github.com/EcoClimLab/ForestGEO-climate-sensitivity/issues/51)**)
*Climwin* runs all potential models to select the best fit (lowest AIC), and does k-fold cross-validation in its computation of AIC to guard against over-fitting [@van_de_pol_identifying_2016]. 
For each group of candidate climate variables, we move forward with the best variable over the time window identified by *climwin* as a candidate climate variable for the multivariate models.
